---
title: "Time series decomposition"
output: html_document
---

  - Show that a 3x5MA is equivalent to a 7-term weighted moving average with weights of 0.067, 0.133, 0.200, 0.200, 0.200, 0.133, and 0.067.

1/3 * [1/5 * (\y_{t-2}\ + \y_{t-2}\ + \y_{t}\ + \y_{t+1}\ + \y_{t+2}\) + 1/5 * (\y_{t-1}\ + \y_{t}\ + \y_{t+1}\ + \y_{t+2}\ + \y_{t+3}\) + 1/5 * (\y_{t}\ + \y_{t+1}\ + \y_{t+2}\ + \y_{t+3}\ + \y_{t+4}\)] 


    plastics dataset
```{r}
autoplot(plastics)
```

  - The plot shows an increasing trend and a seasonal pattern.
```{r}
fit <- plastics %>% decompose(type="multiplicative") 

autoplot(fit) + xlab("Year") +
  ggtitle("Classical multiplicative decomposition
    of a product A for a plastics manufacturer")
```

  - The plots confirm the interpretation of the plot shown earlier.

```{r}
autoplot(plasticsbis, series = 'Data') +
  autolayer(seasadj(fit), series = 'SA CD') +
  autolayer(seasadj(stl(plasticsbis, s.window = 'periodic')), series = 'SA STL') +
  xlab('Year') +
  ylab('Sales (Thousands)') +
  ggtitle('Monthly sales (in thousands) of product A for a plastics manufacturer') +
  scale_color_manual(values = c('black', 'red', 'green'), 
                     breaks = c('Data', 'SA CD', 'SA STL'))
```


  Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?
  
```{r}
set.seed(123)
index <- sample(1:length(plastics), 1)
plasticsbis <- plastics
plasticsbis[index] <- plastics[index] + 500
plasticsbis[index]
plastics[index]
```

```{r}
autoplot(plasticsbis, series = 'Data') +
  autolayer(seasadj(stl(plasticsbis, s.window = 'periodic')), series = 'SA with outlier') +
  xlab('Year') +
  ylab('Sales (Thousands)') +
  ggtitle('Monthly sales (in thousands) of product A for a plastics manufacturer') +
  scale_color_manual(values = c('black', 'red'), 
                     breaks = c('Data', 'SA with outlier'))
```

  - the outlier created a spike in the Seasonally Adjusted data but does little to increase the volatility of the seasonally adjusted data.

  Does it make any difference if the outlier is near the end rather than in the middle of the time series?
```{r}
set.seed(123)
plastics3 <- plastics
plastics3[length(plastics)] <- plastics[length(plastics)] + 500
plastics3[length(plastics3)]
plastics[length(plastics3)]
```  

```{r}
autoplot(plastics3, series = 'Data') +
  autolayer(seasadj(stl(plastics3, s.window = 'periodic')), series = 'SA with outlier') +
  xlab('Year') +
  ylab('Sales (Thousands)') +
  ggtitle('Monthly sales (in thousands) of product A for a plastics manufacturer') +
  scale_color_manual(values = c('black', 'red'), 
                     breaks = c('Data', 'SA with outlier'))
```

  - he outlier at the end does not influence the seasonal component as much as the outlier in the middle. But, the end point of the seasonally adjusted time series is larger than the original data and the trend component is smoother.

  - So the outliers in the middle of time series will have a significant impact on the seasonal component. The outliers at the end will result in a smoother trend.

    cangas dataset
```{r}
autoplot(cangas) +
  ylab('Year') +
  ylab('gas production in billions of cubic meters') +
  ggtitle('Monthly Canadian gas production in billions of cubic meters')

ggsubseriesplot(cangas)
ggseasonplot(cangas)
```

  - The time plot shows a clear increasing trend and a strong seasonality. This confirmed by the subseries plot and the seasonal plot. The gas production increases during winter and decreases in summer. The seasonal plot shows a strong seasonality between 1973 to 1990.
  
  Do an STL decomposition of the data. You will need to choose s.window to allow for the changing shape of the seasonal component.
```{r}
cangas %>%
  stl(t.window = 13, s.window = 15, robust = T) %>%
  autoplot() + ggtitle('STL Decomposition')
```

  - The trend is increasing over time and the seasonality increases around 1974 and startes decreasing a bit in 1990. 
 
  Compare the results with those obtained using SEATS and X11. How are they different? 
```{r}
library(seasonal)
cangas %>% seas(x11="") %>%
  autoplot() +
  ggtitle("X11 decomposition of Monthly Canadian gas production")


cangas %>% seas(x11="") %>%
  autoplot() +
  ggtitle("SEATS decomposition of Monthly Canadian gas production")
```

  - The trend and seasonal components from stl and X11 look similar. The changes of seasonality are different from the original data. This is because the seas() function assumes the multiplicative decomposition; while the stl() function assumes the additive decomposition. 
  
  
  Use stlf() to produce forecasts of the writing series with either method="naive" or method="rwdrift", whichever is most appropriate. Use the lambda argument if you think a Box-Cox transformation is required.

```{r}
p1 <- autoplot(stlf(writing, method='naive')) + ggtitle('Forecast with naive method')
p2 <- autoplot(stlf(writing, method = 'rwdrift')) + ggtitle('Forecast with a drift method')
p3 <- autoplot(stlf(writing, lambda = BoxCox.lambda(writing))) + ggtitle('Forecast with a drift method and Box-Cox transformation')
gridExtra::grid.arrange(p1, p2, p3, ncol = 2, nrow = 2)
```

  - The time series shows an increasing trend and a strong seasonality, so the drift method will outperform the naive method.
  - Applying a Box-Cox transformation yield to a better result (narrower prediction intervals), Because it impose the seasonality to be constant over time. 
  
  Use stlf() to produce forecasts of the fancy series with either method="naive" or method="rwdrift", whichever is most appropriate. Use the lambda argument if you think a Box-Cox transformation is required.

```{r}
p1 <- autoplot(stlf(fancy, method='naive')) + ggtitle('Forecast with naive method')
p2 <- autoplot(stlf(fancy, method = 'rwdrift')) + ggtitle('Forecast with a drift method')
p3 <- autoplot(stlf(fancy, lambda = BoxCox.lambda(fancy))) + ggtitle('Forecast with a drift method and Box-Cox transformation')
gridExtra::grid.arrange(p1, p2, p3, ncol = 2, nrow = 2)
```

  - The forecast using drift method and a Box-Cox transformation yield to abetter results comapred to the naive or a simple drift method.